---
title: "HW1"
author: "Andrew Jowe"
date: "2025-04-15"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


bmatrix_latex <- function(mat) {
  rows <- apply(mat, 1, function(row) paste(row, collapse = " & "))
  body <- paste(rows, collapse = " \\\\ ")
  paste0("\\begin{bmatrix}", body, "\\end{bmatrix}")
}
```

## Problem 1

First we input the data.
```{r}
mu = c(3, 1, 4)
sigma = matrix(c(6, 1, -2, 1, 13, 4, -2, 4, 4),3,3)
```

Here we use the fact that if $X\sim\mathcal{N}_d(\mu,\Sigma)$, for $A\in\mathbb{R}^{k\times d}$, $AX\sim\mathcal{N}_k(A\mu,A\Sigma A^T)$.

### (a)
```{r}
A = t(c(2, -1, 3))
A%*%mu
A%*%sigma%*%t(A)
```

### (b)
```{r}
A = rbind(c(1, 0, 0),c(0, 0, 1))
A%*%mu
A%*%sigma%*%t(A)
```

### (c)
```{r}
A = rbind(c(1, 0, 0),c(0, 0, 1),c(0.5, 0, 0.5))
A%*%mu
A%*%sigma%*%t(A)
```

## Problem 2
In this problem, we utilize the fact that if $X\in\mathbb{R}^m,Y\in\mathbb{R}^n$ and
$$
\begin{pmatrix}
X\\Y
\end{pmatrix}\sim\mathcal{N}_{m+n}\left(
\begin{pmatrix}
\mu_x\\\mu_y
\end{pmatrix},\begin{pmatrix}
\Sigma_{xx},\Sigma_{xy}\\
\Sigma_{yx},\Sigma_{yy}
\end{pmatrix}
\right)
$$
we therefore have 
$$Y|X\sim\mathcal{N}_n(\mu_y+\Sigma_{yx}\Sigma_{xx}^{-1}(X-\mu_x),\Sigma_{yy}-\Sigma_{yx}\Sigma_{xx}^{-1}\Sigma_{xy})$$

This formula is complicated. Please write it in your cheat sheet in exams!

Some rule of thumb to memorize. Since you are actually get the distribution of $Y$, the conditional mean and covariance should be the same dimension as that of $Y$. Look at $\Sigma_{yx}\Sigma_{xx}^{-1}\Sigma_{xy}$, you can think of $x$'s here are cancelled out and remember that our conditional covariance will be reduced so the sign is negative. For the mean part, similarly, in $\Sigma_{yx}\Sigma_{xx}^{-1}(X-\mu_x)$, $x$'s are cancelled out, there is one $x$ in $\Sigma_{yx}$, two reciprocal of $x$ in $\Sigma_{xx}^{-1}$ and one $x$ in $(X-\mu_x)$.


Then we input the data
```{r}
mu_y = c(2, -1)
mu_x = c(3, 1)
sigma_yy = rbind(c(7, 3),c(3, 6))
sigma_yx = matrix(c(-3, 2, 0, 4),2,2)
sigma_xy = t(sigma_yx)
sigma_xx = cbind(c(5, -2),c(-2, 4))

sigma_yy
sigma_yx
sigma_xx
```

Now do the math
```{r}
step_1 <- sigma_yx%*%solve(sigma_xx)
step_1

step_2 <-mu_y-sigma_yx%*%solve(sigma_xx)%*%mu_x
step_2

step_3 <- sigma_yy-sigma_yx%*%solve(sigma_xx)%*%sigma_xy
step_3
```

```{r include=FALSE}
latex_step_1 <- bmatrix_latex(step_1)
latex_step_2 <- bmatrix_latex(step_2)
latex_step_3 <- bmatrix_latex(step_3)
```

So we have

$$\mathbb{E}(Y|X)=`r latex_step_2`+`r latex_step_1`X$$

$$\mathbf{COV}(Y|X) = `r latex_step_3`$$

Our distribution is:

$$y|x \sim \mathcal{N}(\mathbb{E}(Y|X), \mathbf{COV}(Y|X))$$
$$y|x \sim \mathcal{N}(`r latex_step_2`+`r latex_step_1`X, `r latex_step_3`)$$